---
title: "HealthWelfareStatDB"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HealthWelfareStatDB}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(HealthWelfareStatDB)
```

The HealthWelfareStatDB package contains functions for connecting to the web API 
for the Statistical Database at the Swedish National Board of Health and Welfare
(Socialstyrelsen). The documentation for the API can be found 
at: https://sdb.socialstyrelsen.se/sdbapi.aspx. This package has been developed 
to work for version 1 of the API. 

Note: there seem to be a limitation in the API in that it cannot handle
queries for multiple measure categories. For all other variables, it seems to
be possible to query multiple categories in one query.

There is one function to retrieve information on the metadata (return_meta), 
and one function to retrieve data (return_data). Both functions return data
frames.

The function return_meta can be used to query the metadata. This call lists 
the available versions:
```{r}
versions <- return_meta(type="api_version")
print(versions)

```

This call lists the languages available:
```{r}
languages <- return_meta(type="lang")
print(languages)

```

For a selected language, one can view the topics:
```{r}
topics_en <- return_meta(type="topic", lang="en")
print(topics_en)
topics_sv <- return_meta(type="topic", lang="sv")
print(topics_sv)
```

For a selected topic and language, view available variables:
```{r}
vars <- return_meta(type="var", lang="en", topic="diagnoserislutenvard")
print(vars[,1:2])

```

View variable categories, for region:
```{r}
var_cats_region <- return_meta(type="var_cat", lang="en", 
                            topic="diagnoserislutenvard", var="region")
print(var_cats_region)

```

View the first 20 variable categories, for diagnosis. The values in the "kod" 
columns are the ICD-10 codes (International Statistical Classification of 
Diseases and Related Health Problems – Tenth Revision).
```{r}
var_cats_dia <- return_meta(type="var_cat", lang="en", 
                            topic="diagnoserislutenvard", var="diagnos")
names(var_cats_dia)
print(var_cats_dia[1:20,1:3])

```
Data can be queried by entering desired values/categories for the variables, for
a selected language and topic. First, obtain list of variables for the topic of 
choice, here In-Patient Care Diagnoses (diagnoserislutenvard)
```{r}
vars <- return_meta(type="var", lang="en", topic="diagnoserislutenvard")
var_list <- vars[,1]

```

Then construct the variable values list by entering the ids for the variable categories:
```{r}
#  Region: Entire Sweden and Kalmar
#  Age group: 40-44,45-49
#  Sex: Male and female
#  Measure: Number of patients per 100,000 inhabitants
#  Years: 2012 and 2013
#  Diagnosis: J13 and J14 (pneumonia)
values_list <- c("0,8",
                "9,10",
                "1,2",
                "7",
                "2012,2013",
                "J13,J14")
```

Then construct the paired variable-variable category input dataframe

```{r}
df_input_vars <- as.data.frame(cbind(var_list,values_list))

```

Review the input:
```{r}
print(df_input_vars)

```

Send query for the data. The option ```addText=TRUE ``` enables variables with 
text labels to the id variables to be added as ordered factor variables, except 
for diagnosis and year which is left to the user to handle. The ordering follows the
value of the id variable.
```{r}
data <- return_data(lang="en",topic="diagnoserislutenvard", df_input_vars, addText=TRUE)
print(data)
```




